Historical Soil Organic Carbon Budget

### 2020-06-18 SOCbudgetPaper/introduction.Rmd
Update introduction.Rmd

Within the last centuries the recognition of soil organic carbon (SOC) has developed from being a major source of agricultural carbon emissions towards a potentially large additional sink to offset other green house gas emissions and thus mitigate climate change. Whereas the extent of sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, over topping the atmospheric and even more the biospheric carbon pool multiple times. Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref. permafrost melting). The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).

Mapping the worlds SOC has researches intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools. It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard. Nevertheless,
these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for SOC dynamics till dawn of civilization <!--to heroic find another word-->.

Recent studies identified the SOC debt of anthropogenic source at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006). Other studies have focus more closely on spatially disaggregation SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->). Most of these studies doing cutting edge research in their field, but seem to lack proper data on land management.

On the other hand small-scale models (ref. Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities. Due to the lack of proper, concise management data to feed these input demanding models on larger extends, it seems still very complex to scale them up to a global extent.

This study however wants to contribute filling the gap on impacts of agricultural management like yield levels, residue recycling, manure amendments, irrigation and tillage on SOC stocks of the worlds croplands. It is doing so by combining work on agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.

@@

>>>>

<!-- this is not a mitigation paper. Your first paragraph is however on mitigation. I would frame it closer to your research. This is a better intro for your next paper -->

The depletion of Soil organic carbon (SOC) due agricultural management has been a major source of carbon emissions, but also offers the potential to turn into a large additional carbon sink and thus mitigate climate change. Whereas the extent of the sink potential by SOC enhancement is highly debated (add refs), it is largely agreed upon that the SOC pool itself is the biggest terrestrial carbon pool, exceeding the atmospheric and even more the biospheric carbon pool multiple times. Even small changes in SOC drivers might lead to substantial shifts in earth carbon cycled and influence the atmospheric CO2 concentration (ref. permafrost melting). The specific amount of carbon stored in the soil is uncertain though and estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile (Bathjes, 1996).

@@

Mapping the worlds SOC has been researched intensivly over the last decades and led to an increased quality of SOC maps as well as a even better understanding of factors driving magnitude, distribution and dynamics of SOC pools. It is undisputed that natural properties like climatic, biophysical and landscape characteristics play the most important role in this regard. Nevertheless,
these factors can not be easily altered by human intervention, leaving land cover, land use and land management changes the most important factors for current SOC dynamics. <!--to heroic find another word-->.

Recent studies identified the anthropogenic SOC debt at around 116 GtC (Sanderman et al.), which compares to older estimates of around 60-130 GtC <!--look up --> (Lal, 2006). Other studies have focused more closely on spatial disaggregated SOC changes via advanced digital soil mapping techniques (S-World; Stoorvogel 2, 2017) or better representation of biogeochemcial processes within SOC dynamics  (<!-- find microbial study -->). Despite their importance in their field, these studies lack a detailed consideration of land management.

Also small-scale models (ref. Daycent, RothC, Ecosse, C-Tool) are able to represent well SOC dynamics on field-scale driven by detailed management data on yield levels, fertilizer inputs and various other on farming activities. Due to the lack of comprehensive global management data to feed these input demanding models, it seems still very complex to scale them up to a global extent.

Our study therefore estimates the change of global SOC stocks under specific consideration of agricultural mangement, including the impacts of different yield levels, residue recycling, manure amendments, irrigation and tillage. Our approach combines an spatial-explicit estimate of agricultural management data on global level with a robust but light weighted SOC model to estimate SOC stocks and stock change factors as well as organic carbon flow dynamics within the agricultural system.
